### 2015-03-27 paper/plos.tex
address referees comments Tested everything on crystal.

% Template for PLoS
% Version 3.0 December 2014
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended 
% to minimize problems and delays during our production 
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % % 
%
% Once your paper is accepted for publication, 
% PLEASE REMOVE ALL TRACKED CHANGES in this file and leave only
% the final text of your manuscript.
%
% There are no restrictions on package use within the LaTeX files except that 
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% Please do not create a heading level below \subsection. For 3rd level headings, use \paragraph{}.
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file. 
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% See http://www.plosone.org/static/figureGuidelines for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - tabs/spacing/line breaks within cells to alter layout or alignment
% - vertically-merged cells (no tabular environments within tabular environments, do not use \multirow)
% - colors, shading, or graphic objects
% See http://www.plosone.org/static/figureGuidelines#tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://www.plosone.org/static/latexGuidelines
%
% Please be sure to include all portions of an equation in the math environment.
%
% Do not include text that is not math in the math environment. For example , CO2 will be CO\textsubscript{2}.
%
% Please add line breaks to long display equations when possible in order to fit size of the column. 
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% % % % % % % % % % % % % % % % % % % % % % % % 
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% fixltx2e package for \textsubscript
\usepackage{fixltx2e}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb,amsfonts}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% rotating package for sideways tables
\usepackage{rotating}

%Author's added to template:
\usepackage{color}
\newcommand\micron{\ensuremath{\mu\text{m}}}
%%\newcommand{\red}[1]{{\bf \color{red} #1}}
\newcommand{\fixme}[1]{\red{[#1]}}

% Remove comment for double spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2009}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\lhead{\includegraphics[natwidth=1.3in,natheight=0.4in]{PLOSlogo.png}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\sf PLOS}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

%% END MACROS SECTION
\newcommand{\red}[1]{{\bf \color{red} #1}}
\newcommand{\blue}[1]{{\bf \color{blue} #1}}
\newcommand{\green}[1]{{\bf \color{green} #1}}

\begin{document}
\vspace*{0.35in}

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf\newline{Theoretical prediction of disrupted Min oscillation in
  flattened \emph{Escherichia coli}}
}
\newline
% Insert Author names, affiliations and corresponding author email.
\\
Jeff B. Schulte\textsuperscript{1,*},
Rene W. Zeto\textsuperscript{1},
David Roundy\textsuperscript{1}
\\
\bf{1} Dept. of Physics/Oregon State University/Corvallis, Oregon, USA
\\

% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
% 
% Remove or comment out the author notes below if they aren't used.
%

* E-mail: Corresponding schuljef@physics.oregonstate.edu
\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
  The dynamics of the Min-protein system help \emph{Escherichia coli}
  regulate the process of cell division by identifying the center of
  the cell.  While this system usually exhibits robust bipolar
  oscillations in a variety of cell shapes, recent experiments have
  shown that when the cells are mechanically deformed into wide,
  flattened out, irregular shapes, the spatial regularity of these
  oscillations breaks down. We employ widely used stochastic and
  deterministic models of the Min system to simulate cells with
  flattened shapes.  The deterministic model predicts strong bipolar
  oscillations, in contradiction with the experimentally observed
  behavior, while the stochastic model, which is based on the same
  reaction-diffusion equations, does predict spatially irregular
  oscillations as observed in experiment.  We further report
  simulations of symmetric but flattened cell shapes, and find that it
  is the flattening and lateral expansion rather than the asymmetry of
  the cell shapes that causes the irregular oscillation behavior.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step. 
% Author Summary not valid for PLOS ONE submissions.   
\section*{Author Summary}
  When an E. Coli cell divides it must split into two relatively equal
  volumes in order to survive.  One mechanism that aids in this
  process is a system of proteins that oscillate between cell ends and
  help the original cell divide along its center.  A recent
  experiment found that by drastically deforming these cells this
  oscillation can be disrupted.  We examine these oscillations in
  similarly deformed cells computationally, using two widely used
  simulation models, and find good agreement using one of them.  The
  other model fails in a manner far more dramatic than has been
  previously observed.  Furthermore, we find that the effect observed
  experimentally in asymmetric and flattened cells is primarily a
  result of the flattening, not the asymmetry.

\linenumbers

\red{Huh}

\fixme{
  First referee:
  \begin{enumerate}
  \item Done%% \blue{Done} Add discussion of ``platform'' --- wrote our own code,
    %% publically available (cite github?).
  \item Done
    %% \blue{Done} Make 3D geometries available? Cite that the publically
    %% available code can generate them?
  \item Done%%Claims differential equations aren't needed.
  \item Doen%%Claims wild-type movie-ish plot not needed.
  \item Points out problems with Huang 2003:
    \begin{enumerate}
    \item \blue{David.  I've added some sentences discussing the
      starry night effect and membrane diffusion blurring out the
      stars.  I have some confusion however about the diffusion
      lengths.}  Lacks diffusion on the membrane. Supported by Meacci
      2006, which experimentally tries to measure diffusion on
      membrane, has a tough time, but shows that definitely diffuses.
    \item \blue{David} Underestimates diffusion in cytoplasm by a factor of 5-6.5.
      Supported by Meacci 2006, who concludes a rate of 16 microns
      squared per sec for MinD and 10 microns squared per sec for
      MinE.
    \item \blue{David} Uses MinD-ADP -> MinD-ATP that is 3 times lower than
      experimental bound. Halatek uses ADP->ATP of 1/6 s-1. They cite
      Meacci 2006, who state that there is a lower bound on reference
      time of 3 seconds, so this constant has an upper bound of 1/3
      s-1.
    \item \blue{David} (Still debated) No evidence that DE recruits to membrane.
    \item Done%% Works in only a limited part of parameters phase space, and
      %% could not account for important things like
      %% temperature-independence.
    \end{enumerate}
  \item Done%% \blue{Done} Height should be bigger to compare with experiment.  0.4-0.6
    %% microns.  The 0.25 micron height of the cells used in this study
    %% does not represent the experimental settings well. The simulations
    %% are based on the flattened cell shapes reported from the
    %% experiments by Mannik et al 2009 and 2012. In these experimental
    %% studies, E. coli were shown to be able to grow in channels as low
    %% as 0.6 micron in hard silicon structures, and no more than 0.4
    %% micron (see Mannik 2009). Indeed, 0.25 micron channels were
    %% constructed with PDMS ceilings, which is a rather flexible polymer
    %% allowing bacteria to deform it to effectively higher
    %% channels. Thus the simulations would be more realistic if
    %% performed in the range of 0.4-0.6 micron thickness.

    %% \emph{If you can't make it grow in a .3 micron (rigid) silicon channel,
    %%   then we and you observe it growing in a .25 micron flexible
    %%   channel, it presumably must be able to grow because it deformed
    %%   the channel to larger than .3 microns.}
  \item Done%%\blue{David} Starry night in the wild-type cells is a problem.
  \item Done%%Better explanation of correlation function (which is
    %%interesting!).
  \item Done%%Shape A is the interesting bit in our main figure.
  \item Done%%Would like more analysis of stochastic locations in different
    %%geometries.  Clusters?  Could add more geometries?
  \item Done%%Bold conclusions about flattening.  Talk about both width and
    %% thickness.  Compare Halatek 2012.  Wants systematically more
    %% geometries again.
  \item Done%%Is doubtful of our correlation plots relative to those of
    %%Fange and Elf 2006.
  \item Done\blue{David. Read the refferee report with regard to this, and
    not sure what he wants...} Clarify (esp in conclusions) what we
    mean by flattening and irregular shape.  ``breaking the
    cylindrical symmetry and flattening the cells.'' versus ``breaking
    the mirror plane symmetry of the cross-section of a flattned
    cell.''
  \item Done%%[TODO suggestion] Switch to Halatek and Frey
  \item Done%%[TODO suggestion] Thoroughly explor shape parameter space
  \item Done%%[TODO suggestion] Check robustness of our conclusions with
    %%respect to changes in (reaction-diffusion?) parameters.
  \item Done%%Check order of figures and their references.
  \item Done%%%% \blue{David.  I cited the Corbin paper (although he's got a
    %% typo because Corbin 2008 is after the fange 2006 paper. I could
    %% cite other experimental papers, but the sentence in question
    %% specifically refers to changes made to the Huang model, and the
    %% papers cited are the main reasons that those changes have been
    %% made.)} Cite biological (non-computational) works, that the
    %% e.g. Corbin 2008.
  \end{enumerate}
  Second referee:
  \begin{enumerate}
  \item Done%%Nitpicks about language of deterministic model versus
    stochastic model.  (Read carefully)
  \item Done%%\blue{David} Explain in greater detail the obviousness (maybe) that the
    %% deterministic approach doesn't have fluctuations.  RESPOND:
    %% chaotic systems do deterministically result in fluctuations.
    %% Explain this somewhere, and discuss triviality of the conclusion
    %% that deterministic is wrong because it does not exhibit these
    %% irregularities.
  \item Would like temporal averages over cell-cycle time scales.
  \item More discussion of how flattening leads to spatial
    irregularities.  Compare with large nearly spherical e. coli Shih
    2005.  See also Halatek and Frey 2012 for rotating oscillation
    axis due to destabilization of short modes.  How does this connect
    with our stability analysis?
  \item Done%%\blue{David} Should mention more recent revisions of the model.
  \item Done%%\blue{David} Fange and Huang do not use the same reaction-diffusion
    %%equations.
  \item Done%%%% \blue{David.  I read the referee's note about this.  Not sure
    %% what he/she wants.}  Clarify statement about 1000 proteins and
    %% stochastic fluctuation.  Reword about a continuum description
    %% omitting information?
  \item Done%%%% \blue{Done} Do not call Huang parameters ``wild type values.''  They are
    %% fitted parameters.
  \item Done%%\blue{Done?  I think our additions to the Methods section
    %% addressed this?}  Add methods section for numerics
    %% (deterministic).
  \item Done%%\blue{Done} Add details of the linear stability analysis.
  \item Done%%\blue{David} Still need to fix periods on correlation plots.
    %% \blue{About this point - I updated the data and it seems the
    %%   result is defferent from before - the stadium shapes seem to be
    %%   temporally correlated for much longer compared with the shape A
    %%   than before.}
  \end{enumerate}
}

\section*{Introduction}
It is vital that during the process of bacterial cell
division a cell avoid minicelling, or splitting into daughter cells
with lopsided volumes.  Instrumental to this process in
\emph{Escherichia coli} is a long FtsZ polymer chain that develops on
the cell wall in the center region of the cell, helping dictate the
plane of division~\cite{adams2009bacterial,
  lutkenhaus2007assembly}. Previous experimental studies have shown
that the MinC protein, known to inhibit the FtsZ
polymer~\cite{shen2010examination}, exhibits regular pole to pole
oscillatory behavior between both ends of the wild-type pill-shaped
cell.  It thus has a higher time averaged concentration in the cell
poles than in the center region, which aides in preventing the FtsZ
from developing in the wrong region.  The MinC is recruited to these
poles by MinD, which itself interacts with another protein, MinE, in a
system which exhibits pole-to-pole oscillatory
behavior~\cite{shapiro2009and, yu1999ftsz,
  meacci2005min,huang2003dynamic,kerr2006division,mannik2009bacterial}.
%% \cite{hu1999topological, fu2001mine, shapiro2009and,
%%   yu1999ftsz, raskin1999rapid, meacci2005min, raskin1999minde}.

Previous experimental studies have shown that the MinD protein system
is capable of exhibiting oscillations in round
shapes~\cite{fange2006noise} as well as in connected three pronged
tube shapes~\cite{varma2008min}, in which the oscillations seem to
seek out the extreme poles in the cell.  One conclusion of these
studies has been that the MinD system robustly forms bipolar
oscillations regardless of variations in cellular shape.

However, Mannik \emph{et al.} have recently shown that there are
limitations to this robust capability to
oscillate~\cite{mannik2012robustness,mannik2009bacterial,mannik2010bacteria}. They
have experimentally forced \emph{E. coli} cells into microfabricated
channels of thickness less than that of the natural diameter of the
cells, in which they area able to grow and divide. Upon entering the
channels, the cells undergo a mechanical deformation in which they
both widen and lengthen within the plane of the channel.  This
deformation results in very wide (they can reach widths of over
$5\micron$) flattened cells that when viewed from the top down have
irregular and asymmetric shapes.  While these cells are still able to
divide into surprisingly equal volumes, the MinD oscillations in these
cells are irregular, both temporally and spatially. Seen with
fluorescent microscopy, the MinD maximize in multiple locations within
the cell, in a seemingly random sequence. These experiments allow for
an opportunity to test MinD simulation models against more extreme
cases than have been seen thus far.  An additional interesting
question is whether the irregularity observed by Mannik \emph{et al.}
indicates that the MinD system can become chaotic (as in
e.g. Ref.~\citenum{kuramoto1978diffusion}), or if this irregularity is
the result of a stochastic but nonchaotic process.
%
A number of models of the MinD protein system have been developed that
accurately describe its basic oscillatory nature.
%
Early models involved free proteins that affect each others' rates of
diffusion and membrane attachment, but do not combine into compound
states~\cite{meinhardt2001pattern}.  In 2003 Huang improved upon this
approach with a simple and very successful simulation model based on
MinD-MinE combination, ATPase hydrolysis, and MinD membrane attachment
that exhibits accurate MinD oscillations in cylindrical
cells~\cite{huang2003dynamic}. In this model cytoplasmic MinD is
recruited to the membrane by MinD that is already clustered there
(following observed non-linear attachment of MinD on the cell
membrane~\cite{hu2002dynamic,shih2002division}), and is stationary
once attached.
%

\blue{5a,b,c,d}Several models
\cite{fange2006noise,bonny2013membrane,halatek2012highly} modify
Huang's original model to more accurately reflect experimental
findings, which were published after his 2003
paper~\cite{meacci2006mobility,loose2011min}. These models introduce
diffusion along the membrane that is two to three orders of magnitude
slower than that within the cytoplasm, and increase the rate of
cytoplasmic diffusion by a factor of about five.  They also increase
or assume instantaneous~\cite{bonny2013membrane} the nucleotide
exchange rate.  In addition, they revise the process of MinD
recruitment to the membrane, prohibiting MinD from attaching to
MinD-MinE complexes there, behavior that is experimentally unresolved.
The model of Bonny \emph{et al.} allows MinE to remain independently
attached to the membrane after dissasociating with
MinD~\cite{bonny2013membrane}.
%

There are only on the order of 1000 MinD proteins in a given cell,
which is significantly less than the number of gridpoints used in our
simulations, suggesting that a continuum description will break down
and stochastic variation plays a significant role.  To treat this,
there have been a number of variations of the Huang 2003 model that
stochastically simulate the similar reaction-diffusion
equations~\cite{fange2006noise, kerr2006division}.  These studies
largely confirm the results of Huang's deterministic model when
applied to the wild-type, pill-shaped phenotype.  However, the
stochastic models are slightly more successful in predicting
experimentally observed oscillations in round cell
phenotypes~\cite{fange2006noise, huang2004min}, and they enable
prediction of fluctuations in the predicted
behavior~\cite{kruse2007experimentalist}.  Both deterministic and
stochastic models are widely used throughout systems
biology~\cite{lawson2013spatial, robb2014stochastic,
  oguz2014stochastic, fu2013deterministic, rudiger2014stochastic}, and
have unique advantages and limitations.  The deterministic approach
has an advantage in providing a simple prediction of average behavior,
while the stochastic approach enables prediction of fluctuations from
that mean, and reduces sensitivity to initial conditions.

%% biology~\cite{lawson2013spatial, robb2014stochastic,
%%   oguz2014stochastic, fu2013deterministic, rudiger2014stochastic}, and
In this paper we use Huang's original 2003
model~\cite{huang2003dynamic}, which has been widely
used~\cite{hattne2005stochastic,huang2004min,kerr2006division} in both
deterministic and stochastic variants, to study flattened cells of
$0.4\micron$ thickness that are similar to those observed by Mannik
\emph{et al.}~\cite{mannik2009bacterial}.

\begin{figure}
  \includegraphics[width=8.7cm]{reactions}
  \caption{Reactions included in the model of Huang \emph{et
      al.}~\cite{huang2003dynamic}.}\label{fig:reactions}
\end{figure}

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section*{Materials and Methods}
\subsection*{Model and Cell Shapes}
\label{sec:model-method-shapes}
We implement the reaction-diffusion model of Huang \emph{et
  al.}~\cite{huang2003dynamic}.  Figure~\ref{fig:reactions} shows the
reaction process.  The cytoplasmic MinD:ADP complex undergoes
nucleotide exchange and is changed into the MinD:ATP complex.  This
will naturally diffuse and attach to the cell membrane.  A cytoplasmic
MinE will attach to the wall bound MinD:ATP complex and after a time
will activate ATP hydrolysis.  This breaks up the complex, releasing
MinE, phosphate, and MinD:ADP back into the cytoplasm.  The MinD:ADP
will undergo nucleotide exchange and begin again the cyclic process.
The model is defined by a set of five reaction-diffusion equations:

\begin{align}
  \frac{\partial \rho_{D:ADP}}{\partial t} = \mathcal{D}_D\nabla^2\rho_{D:ADP}-k_D^{ADP\rightarrow ATP}\rho_{D:ADP}+\delta(d_w)k_{de}\sigma_{DE},\hspace{3.4cm}
\end{align}
\begin{align}
  \frac{\partial \rho_{D:ATP}}{\partial t} = \mathcal{D}_D\nabla^2\rho_{D:ATP}+k_D^{ADP\rightarrow ATP}\rho_{D:ADP}-\delta(d_w)[k_D+k_{dD}(\sigma_D+\sigma_{DE})]\rho_{D:ATP}
\end{align}
\begin{align}
  \frac{\partial \rho_E}{\partial t} = \mathcal{D}_E\nabla^2\rho_E+\delta(d_w)k_{de}\sigma_{DE}
  -\delta(d_w)k_E \sigma_D \rho_E
\end{align}
\begin{align}
  \frac{\partial \sigma_D}{\partial t} &= -k_E\sigma_D\rho_E
  +[k_D+k_{dD}(\sigma_D+\sigma_{DE})]\rho_{D:ATP}
  \label{eq:d-on-wall}
\end{align}
\begin{align}
  \frac{\partial \sigma_{DE}}{\partial t} &= -k_{de}\sigma_{DE}+k_E\sigma_D\rho_E\hspace{3cm}
  \label{eq:FifthPDE}
\end{align}
where $\rho$ is cytoplasmic protein density (proteins$/\micron^{3}$), $\sigma$
is membrane bound density (proteins$/\micron^{2}$), $\mathcal{D}_D$ and
$\mathcal{D}_{E}$ are the diffusion constants for MinD and MinE,
respectively, $k_D^{\textrm{ADP $\rightarrow$ ATP}}$ is the rate of
conversion from MinD:ADP to the MinD:ATP complex, $k_D$ is the rate of
MinD:ATP attachment to the membrane when no protein is already
attached there, $k_{dD}$ is the increase of this rate when MinD:ATP is
present on the membrane, $k_{de}$ is the rate of hydrolysis of the
MinD:MinE:ATP complex, $k_E$ is the rate of cytoplasmic MinE binding
to membrane bound MinD:ATP complex, and $d_w$ is the distance from the
point in space to the closest wall.  The Dirac delta function
$\delta(d_w)$, which we need to describe the location of the membrane,
has units of $\micron^{-1}$ and is zero everywhere except at the wall.
Equations \ref{eq:d-on-wall} and \ref{eq:FifthPDE} are only relevant
at the membrane because the membrane-bound density values have no
meaning in the cytoplasm.

Our diffusion and reaction rates are shown below.  We are interested
primarily in the effect of cellular size and shape on the protein
oscillations, so we use Huang's parameter
values~\cite{huang2003dynamic}.

\begin{gather*}
  \mathcal{D}_D = \mathcal{D}_{E} = 2.5\micron^2/\text{sec}\\
  k_D^{\textrm{ADP $\rightarrow$ ATP}} = 1/\textrm{sec,  }
  k_D = 0.025 \micron /\textrm{sec}\\
  k_{dD} = 0.0015 \micron^3/ \textrm{sec,  }
  k_{de} = 0.7/\textrm{sec}\\
  k_E = 0.093 \micron^3 /\textrm{sec}.
\end{gather*}

Huang's simulations use total MinD and MinE concentrations of
$1,000/\micron$ and $350/\micron$, respectively, in a cylindrical cell
of radius $0.5\micron$, and in our (non-cylindrical) cells we use the
same number of proteins per unit volume.  These concentration values
are $1273\micron^{-3}$ and $446\micron^{-3}$, respectively. We have
written our own simulation platform and membrane creation tools from
scratch, instead of using existing available software used in previous
studies~\cite{bonny2013membrane,fange2006noise,halatek2012highly,kerr2006division}
(we believe there is value in reproducing results with alternative
software).  Our simulations take place within a three-dimensional
Cartesian grid that has a grid spacing of .05\micron, and our cell
shapes are specified as the zero of an analytic function which can
generate the geometry presented in this paper.  Our code is publically
available on github~\cite{git}, including scripts that generate the
simulation data used in this paper.

We have performed both a numerical, deterministic model simulation
that is spatially and temporally discrete, and a stochastic simulation
that is spatially discrete but continuous in time.  Our stochastic
model follows the work of Kraus~\cite{kraus1996crosstalk} which in
turn follows a method introduced by
Gillespie~\cite{gillespie1977exact}.

We mean to investigate the geometric limits of the Min system
oscillations as observed by Mannik \emph{et
  al.}~\cite{mannik2012robustness}, so we have modeled the Min system
in several cell shapes and sizes.  Here we present a selection of
these, beginning with naturally occurring pill-shaped cells, followed
by a number of flattened out shapes which reflect the experiments of
Mannik \emph{et al.}, in which bacteria are confined within a thin
slits.  These slits are fabricated with a width of .25\micron, but
they are coated with a PDMS lining that the cells are able to deform,
raising uncertainty in the actual cell thickness.  Tests with pure
silicon, non-deformable slits show that the greatest thickness for
which cells are able to penetrate is
.4\micron~\cite{mannik2012robustness,mannik2009bacterial}.  We
therefore assume that the PDMS slits have been deformed to this
thickness and simulate flattened cells with a thickness of .4\micron.
Viewed from the top down the cells have the shapes described below and
viewed from the side they have at their edges a semicircular
cross-section (one may imagine the shape of a pancake).
%
In this paper we focus on four specific flattened cell shapes.  Two of
these shapes replicate those published by Mannik \emph{et al.}, and
the other two are `stadium' shapes that respectively have the same
aspect ratio, thickness, and volume as the two cell shapes
experimentally observed by Mannik \emph{et
  al.}~\cite{mannik2012robustness}.  Viewed from the top down, these
stadium shapes appear as rectangles with semi-circular end caps on the
long axis ends.

\begin{figure*}
  %\includegraphics[width=\textwidth]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-exact/plots/image-plot}
  \begin{center}
    \hspace{-3.5cm}
    \includegraphics[height=6.4cm]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-exact/plots/single-image-plot}\\
    \vspace{-1.5em}
    \hspace{-3.5cm}
    \includegraphics[height=6.4cm]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-full_array/plots/single-image-plot}
    \vspace{-1.5em}
  \end{center}
  \caption{Images of the concentration of each protein species in a
    natural pill-shaped bacterium at one-second time intervals. The
    upper plots shows results from the deterministic model and the
    lower shows results from the stochastic model.  The order of
    frames is such that individual MinD proteins begin at the bottom
    of the plot (in the MinD:ATP state in the cytoplasm), and progress
    upward until they reach the MinE:MinD:ATP membrane-bound complex.
    At that point, they will spontaneously dissociate into cytoplasmic
    MinE (the top row) and the starting state of cytoplasmic
    MinD:ADP.}.
  \label{image-p}
\end{figure*}

Huang \emph{et al.}~\cite{huang2003dynamic} has performed a linear
stability analysis on a cylindrical model which shows an upper limit
on a steady state solution of a $2\micron$ half wavelength.  Cells
with dimensions longer than this spontaneously develop spatial
oscillations, while cells that are shorter relax into a homogeneous
steady state.  We have performed a similar linear stability analysis
on an infinite slab with a thickness equal to that of our flattened
cells ($0.40\micron$).  We first solve the model's five differential
equations for the steady state solution, under the constraint that the
total number MinD and MinE molecules matches our simulation.  We write
down as a matrix the linear response of the time derivatives of the
density to small, spatially harmonic density perturbations around the
steady-state density.  The system is stable at this wavelength
provided all eigenvalues of the matrix are negative, indicating that
all perturbations with this wavelength will decay.  We find the
stability limit by searching for the largest wavelength at which the
system is stable.  Through this method we arrive at an upper half
wavelength stability limit of $1.56\micron$ for our $0.40\micron$
flattened cells. As expected, when decreasing the lengths and widths
of our simulated flattened cells so that the longest distance across
the cell is less than this length, the cells stop exhibiting any
oscillatory behavior.  The deterministic model relaxes into a
motionless state and the stochastic model exhibits random fluctuations
without spatial oscillations.

% Results and Discussion can be combined.
\section*{Results and Discussion}
\subsection*{Naturally Occurring Pill Shaped Cells}

\begin{figure}
  \begin{center}
    \includegraphics[height=6.4cm]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-full_array/plots/correlation.pdf}
  \end{center}
  \caption{Temporal correlation function of the total MinD found in
    two opposite polar regions of the pill-shaped cell, shown against
    the correlation time.  Data for both the the deterministic and
    stochastic models are shown.  The stochastic model shows an
    oscillation period of 39.5 seconds and coherence time of 307 seconds.
    The correlation functions are scaled to have the same initial
    value.}
  \label{corr-pill}
\end{figure}

We begin with the naturally occurring pill cell shape.  We piece this
shape together as a cylinder with hemispherical end caps.  This shape
follows the early simulations of Huang \emph{et al.} but differs in
that we have added the end caps for a more natural shape, expecting
similar results.

Figure~\ref{image-p} shows a series of color plots of the density of
proteins at each stage of the reaction cycle. Deterministic simulation
data is shown above and stochastic simulation data is shown below.
The cells shown are $4\micron$ in length, measured from end to end.
We have `smeared out' the stochastic concentrations in a manner meant
to reproduce the images shown by diffraction limited fluorescence
microscopy.  We do so using the two dimensional Gaussian approximation
developed by Zhang \emph{et al.}~\cite{zhang2007gaussian}.  In this
approximation we use a numerical aperture value of 1.3, which is the
same as used by Mannik experimentally, and a wavelength of $650nm$.
Each frame is 2.5 seconds ahead of the last, and each image shows the
concentration of a given state of protein (of the five described in
the reaction model) summed over the coordinate normal to the page.

Figure~\ref{image-p} begins about 300 seconds into the simulation and
shows one period of oscillation.  At $t=0$ there is a high
concentration of MinD:ATP that has accumulated on the membrane at the
bottom of the cell. An important aspect of Huang's model is that the
MinD is attracted to and sticks to the membrane nonlinearly: as it
accumulates there it begins to `recruit' other MinD that is diffusing
in the cytoplasm nearby, causing peaks in concentration to build up on
the walls.  Meanwhile, the MinE creeps downward as it reacts with the
membrane-bound MinD, forming the MinD:MinE:ATP complex, then breaks it
apart and diffuses downward a bit more before it again reacts with
membrane-bound MinD.  This process can be seen in the form of the
well-known ``MinE rings'' (actually, MinE bound to MinD on the
membrane).  These rings are visible in the deterministic model plots
as a green band on the walls from 0 seconds to 4 seconds (and later
from 15 to 23 seconds).  The appearance of these rings in the
deterministic model along side their less obvious appearance in the
stochastic model highlights an advantage of the deterministic model:
the idealization of deterministic data allows one to see patterns in
the averaged behavior that might otherwise be missed.  In the
stochastic model the MinE still exhibits higher concentrations in the
same regions throughout the process, but what would ideally be a
``ring'' pattern is instead an asymmetric collection of maxima, which
would become a ring after phase-locked averaging.

During the formation of these rings, cytoplasmic MinE is diffusing in
the upper portion of the cell and will naturally progress downward,
where there is membrane-bound MinD to react with, leading to a
depletion of MinE in the upper portion of the cell.  As the MinE ring
converges upon the lower end of the cell, MinD that has been released
is able to diffuse upward, past the ring, while still in its MinD:ADP
state and unable to bind to the membrane.  After it undergoes
nucleotide exchange, resulting in MinD:ATP, it is ready to accumulate
on the walls in at the top of the cell, where the MinE has been
depleted.  This can be seen in seconds 10 through 20 in both models,
followed by the subsequent MinE ring formation and movement upward
(beginning the same process in the opposite direction) that can be
seen in seconds 15 through 23.

\blue{(5a)} Both Fig.~\ref{image-p} and its associated movies (see
supplementary material), show that the stochastic model exhibits a
``starry night'' effect characterized by spatially fixed points of
protein density build up, as recruitment leads to clusters of MinD
forming on the wall and then subsequently dissipating.  This results
from the lack of inclusion of diffusion along the membrane in Huang's
model, which would allow these recruitment clusters to spread out.
Experiments have confirmed that diffusion along the membrane does
occur, albeit with a rate two orders of magnitude lower than that in
the cytoplasm~\cite{meacci2005min}.  The `stars' in the effect
typically last for around 10 seconds. The membrane-bound proteins will
travel a distance of around one micron during one of these periods of
time, enough to spread out across the polar section of the one micron
diameter wild-type pill cells.  As we see below, the flattened,
irregular cells are large enough so that the one micron blur would not
completely eliminate the `starry night' variation in maxima position
effect.

%% \fixme{calculating
  %% with $<x^2> = Dt$ and $D_{membrane} = .16\mu^2/s m$, which is just
  %% the experimental cytoplasmic value divided by 100, yields 1.2
  %% microns, which maybe is too high for what we're trying to say?
  %% Halatek uses $D_{membrane} = .013\mu^2/s m$, for which the same
  %% calculation would give .36 microns, which is better.  Halatek
  %% doesn't say why he uses a value that seems to be different than what
  %% Meacci claim to have found.  Halatek just says it's experimentally
  %% valid and cites their work.  What to do?}

We also analyze the temporal periodicity of the system, which we
quantify with temporal correlation function of the total MinD found in
two opposite polar regions.  This correlation function, which is
displayed in Fig.~\ref{corr-pill}, is given by
\begin{equation}
  C(\tau) \propto \int
  (N_{\textit{top}}(t) - \bar N_{\textit{top}})
  (N_{\textit{bottom}}(t+\tau) - \bar N_{\textit{bottom}})dt
\end{equation}
where $N_{\textit{top}}(t)$ and $N_{\textit{bottom}}(t)$ are the total
MinD proteins in the top and bottom thirds of a cell.  These plots
help us in studying the periods and the regularity and stability of
oscillations.  These curves are well fit with a simple decoherence
model, given by the equation
\begin{equation}
  C(\tau) = -\cos\left(\frac{2\pi\tau}{T}\right) e^{-\frac{\tau}{\tau_c}}
\end{equation}
where $T$ is the period and $\tau_c$ is the coherence time.  In the
case of the stochastic model the pill exhibits coherence times of
about 8 periods and in the case of the deterministic model oscillate
indefinitely with complete coherence, indicating that its behavior is
perfectly periodic.

\begin{figure*}
  \centering
  \hspace{-3cm}
  \vspace{1cm}
  \includegraphics[width=16cm]{../paper/plot-ave}
  \caption{We display here arrows depicting successive maxima in space
    and and time overlayed on a color plot of the total MinD density
    averaged over the same time period.  The simulation time covered
    for the flattened cells is 350 seconds, which is the same
    period of time depicted in the experimental data plots of Mannik
    \emph{et al.}~\cite{mannik2012robustness}.  For the wild-type pill
    shape, we only cover 250 seconds, in order to provide a
    useful comparison due to its shorter oscillation period.  The top
    row shows plots published by Mannik \emph{et al.} of the MinD maxima behavior
    and the bottom two rows show our simulations using the stochastic
    and deterministic models, respectively.  We simulated
    approximations to the two shapes observed by Mannik, which we call
    \emph{shape A} and \emph{shape B}.  In addition, we studied two
    flattened stadium shapes which we call \emph{stadium A} and
    \emph{stadium B} corresponding in aspect ratio and thickness to
    the two experimental shapes.  The spatial length scale of all the
    figures shown was identical. Each of the flattened cell shapes
    uses the same color scale for the number of proteins per unit
    area.  Finally, we display the natural pill shape, which was also
    featured in Figs.~\ref{image-p} and~\ref{corr-pill}, with a
    different color scale to reflect the thicker cell containing more
    proteins per cross-sectional area.  }
  \label{randst-plot-ave}
\end{figure*}

%% \begin{figure}
%%   \begin{center}
%%     \includegraphics[height=6.4cm]{../data/shape-randst/0_40-18_50-18_50-95_00-15_00-full_array/plots/correlation.pdf}\\
%%     \includegraphics[height=6.4cm]{../data/shape-stad/0_40-2_35-1_32-0_00-15_00-full_array/plots/correlation.pdf}
%%   \end{center}
%%   \caption{Temporal correlation function of the total MinD found in
%%     two opposite polar regions of the shape \emph{A} (above) and the
%%     stadium \emph{A} (below) cell shapes, shown against the
%%     correlation time.  Data for both the the determinisitic and
%%     stochastic models are shown, and the correlation functions are
%%     scaled to have the same initial value.  For shape \emph{A}, the
%%     stochastic model shows an oscillation period of 52 seconds and
%%     coherence time of 248 seconds, so that it takes roughly 4.8
%%     periods for the behavior to decohere.  For the stadium \emph{A},
%%     the same model shows an oscillation period of 48.4 seconds and
%%     coherence time of 634 seconds, so that it takes roughly 13
%%     periods for the behavior to decohere.}
%%   \label{fig:corr-pancake-A}
%% \end{figure}

\begin{figure}
  \begin{center}
    \includegraphics[width=8.7cm]{../data/shape-randst/0_40-18_60-28_60-94_00-15_00-full_array/plots/correlation.pdf}
    \includegraphics[width=8.7cm]{../data/shape-stad/0_40-2_92-1_18-0_00-15_00-full_array/plots/correlation.pdf}
  \end{center}
  \caption{Temporal correlation function of the total MinD found in
    two opposite polar regions of the shape \emph{B} (above) and the
    stadium \emph{B} (below) cell shapes, shown against the
    correlation time.  Data for both the the determinisitic and
    stochastic models are shown, and the correlation functions are
    scaled to have the same initial value.  For shape \emph{B}, the
    stochastic model shows an oscillation period of 58 seconds and
    coherence time of 289 seconds, so that it takes roughly 5 periods
    for the behavior to decohere.  For the stadium \emph{B}, the same
    model shows an oscillation period of 51 seconds and coherence time
    of 442 seconds, so that it takes roughly 8.6 periods for the
    behavior to decohere.}
  \label{fig:corr-B}
\end{figure}

\subsection*{Comparison of Experimental and Stadium Shapes}
%\label{sec:model-method-shapes}
As described in Section~\ref{sec:model-method-shapes}, \fixme{this
  link?} we will focus on simulation of four illustrative flattened
cell shapes: two shapes created to replicate those shapes observed
experimentally by Mannik \emph{et al.}~\cite{mannik2012robustness}
(shape~\emph{A} and shape~\emph{B}), and for comparison two `stadium'
shapes that have the identical aspect ratio, thickness and volume
(stadium \emph{A} and stadium \emph{B}), shown in
Figure~\ref{randst-plot-ave}.  All of the flattened cells have lost
the rotational symmetry of the orginal pill shapes, but while
shape~\emph{A} and shape~\emph{B} have only one mirror plane symmetry
(that is normal to the flattened plane), stadium~\emph{A} and
stadium~\emph{B} have three.  This allows us to distinguish between
the effect of flattening the cell and the cell shape's irregularity
and asymmetry.  For each of these shapes (in addition to the wild-type
pill shape discussed previously) we have simulated in excess of 15000
seconds (over 4 hours) of evolution of the MinD system using the
deterministic version of Huang's model~\cite{huang2003dynamic}, and in
excess of 100000 seconds (over 27 hours) using the stochastic model.

From each of these flattened simulations, we have chosen a typical 350
second segment to compare with the published results of
Mannik \emph{et al.}~\cite{mannik2012robustness}, exemplified in arrow plots in
which the arrow heads show the location of sequential MinD maxima
within the cell (in Fig.~\ref{randst-plot-ave}).  In addition to
arrows between successive maxima in space and time, we plot as a
colored background the density of MinD proteins averaged over the same
time period.  We note that we have manually verified that our
(computer-generated) arrow plots also reflect a human interpretation
of a movie of the same data.  Finally, for comparison we present the
same plot for a wild-type cell, with a time period of 250 seconds to
account for its short period of oscillation.

In every case, including the wild-type pill shape, we see irregularity
in the location of the maxima when using the stochastic model.  The
deterministic model shows uniformly bipolar oscillation.  We conclude
that the model is \emph{not} chaotic (which would show irregularity in
the deterministic computation), but rather that the irregularity
results purely from stochastic processes.  From these results and how
they compare with experiment, we conclude that the deterministic model
is inadequate to explain experimental observations of the locations of
density maxima of the MinD protein.

%% We found that the deterministic simulations of the MinD system results
%% in robust and regular oscillatory behavoir of polar selection and
%% oscillation in not only symmetrical shapes, such as the wildtype pill
%% shape, but also in very assymetrical, flatttened cells, such as those
%% observed by Mannik.  We also examined larger and smaller versions of
%% these shapes, and found this behavior to be very robust, from the
%% minimum size to sustain oscillation (around \fixme{???} $\mu$m) up to
%% around twice the cell size reported by Mannik.  At larger sizes than
%% this, less regular oscillations occur using the deterministic method,
%% but we note that the cells reported by Mannik are already considerably
%% larger in volume than typical wild type
%% cells~\cite{kubitschek1990cell,kubitschek1968linear,mannik2012robustness}.

The predictions of the stochastic model are roughly similar to the
experimental results in terms of irregularity of the locations of
maxima, although there are locations on the edges of the cells where
experiment shows maxima occurring that we never see in our simulations,
suggesting that the model does not precisely reflect the experiment.
We also note that the stadium shapes in Fig.~\ref{randst-plot-ave}
appear to have maxima location irregularity that is qualitatively
similar to our irregular shapes and to experiment.

However Fig.~\ref{randst-plot-ave} leaves unclear the importance of
the flattening and accompying lateral expansion of the cell versus
cell shape irregularity and asymmetry in developing an understanding
of the origin of the experimentally observed behavior.  Although in
the movies the stadiums appear to display, on average over long time
scales, a somewhat more regular oscillation than appears in the
irregular shapes \emph{A} and \emph{B}, the limited time range of the
arrow plots (350 seconds) is inadequate to make this comparison.  We
therefore return to the correlation function between the number of
proteins in a segment at the top of the cell and the number of
proteins in a segment at the bottom of the cell.
Fig.~\ref{fig:corr-B} shows this plot for shape \emph{B} and stadium
\emph{B}.  E. Coli cell doubling times vary widely depending on
environmental conditions, but will often be between 20 minutes to 100
minutes~\cite{pierucci1972chromosome}, while fluorescent microscopy
experiments have shown that the FtsZ ring builds itself with a half
life of roughly 30 seconds~\cite{stricker2002rapid}.  Here we plot
correlations of correlation times up to 500 seconds, a biologically
relevant timescale.

We see that both in the shape \emph{B} and stadium \emph{B}, the
deterministic model is perfectly periodic.  Both the deterministic and
stochastic model correlations have the same number of proteins in
their cells, and have been normalized with the same normalization
factor, so that they can be directly compared against each other.  The
stochastic simulation of shape \emph{B} has a coherence time of 5
periods, while the corresponding simulation of stadium \emph{B} has a
coherence time of 8.6 periods.
%
The correlation function for stadium \emph{A} is similar to that of
stadium~\emph{B}, with similar coherence time (13 periods, not
shown), while shape~\emph{A} also exhibits a similar coherence time (4.9
periods).
%

From the stochastic correlation data, we conclude that irregularity of
shape has only a moderate affect on the temporal regularity of the
oscillatory behavior, with even the very distorted shape~\emph{B}
exhibiting only a factor of two greater decoherence.

\section*{Conclusion}
We find agreement between our stochastic simulations and experiments
showing disrupted bipolar MinD oscillation in asymmetric flattened
E. coli cells~\cite{mannik2012robustness}.  As observed
experimentally, our simulations predict that MinD maxima will form in
a spatially irregular sequence of locations.  This result builds upon
existing results showing that this model and its variants are
effective in a variety of wild-type and mutated cell
shapes~\cite{fange2006noise, varma2008min, kruse2007experimentalist},
and reinforces that stochastic variations of Huang's 2003
model~\cite{fange2006noise, kerr2006division} has considerable
predictive power beyond the standard wild-type cell.  \fixme{Discuss
  which of these are likely to map to more recent models.}

%
%% We have simulated flattened cell shapes similar to those observed
%% experimentally by Mannik \emph{et al.}, with a stochastic and a
%% deterministic version of Huang's widely used reaction-diffusion model.
%% %
%% We find that the stochastic model predicts experimentally accurate
%% qualitative behavoir in both the wild-type, pill-shaped cells (as
%% found in previous studies~\cite{fange2006noise, varma2008min,
%%   kruse2007experimentalist}), as well as in the new, flattened,
%% irregularly shaped cells.  These flattened cells invariably show
%% spatially irregular formation of maxima as has been observed
%% experimentally.

In contrast to the stochastic model, the original deterministic
version of the model~\cite{huang2003dynamic} predicts behavior that
contradicts experimental observations.  Specifically, the
deterministic model predicts robust and periodic bipolar oscillation
in irregularly shaped cells.  Thus we conclude that this method is
inconsistent with experiment, and cannot be relied upon for
predictions of the behavior of the MinD system.  This is the most
clear example of the inadequacy of the deterministic simulation method
in reproducing experimental behavior of the MinD system to date.
Until now the results of deterministic and stochastic simulation have
largely coincided, with the stochastic method showing minor
differences in behavior when compared to deterministic
models~\cite{kerr2006division, fange2006noise, huang2004min,
  kruse2007experimentalist}.  The regular oscillations in the
deterministic case also allow us to conclude that the model is
non-chaotic.  Because of their similar construction, we suspect that
recent variants of the
model~\cite{bonny2013membrane,fange2006noise,halatek2012highly} would
also prove to be non-chaotic.

We find that flattened but regular and symmetric cells exhibit MinD
oscillations that are qualitatively similar to the oscillations
observed experimentally (and in our simulation) in irregular and
asymmetric cells.  This demonstrates that asymmetry is not required in
order to induce spatially irregular MinD oscillations.  In addition,
the temporal regularity of oscillations is only moderately affected by
irregularity in the shape of flattened cells.  Thus we conclude that
it is the flattening of the cells, and the accompying expansion in the
tranverse direction, rather than their irregular shape that primarily
causes the disruption of regular bipolar oscillation which is observed
by Mannik \emph{et al.}~\cite{mannik2012robustness}.

\section*{Supporting Information}

% Include only the SI item label in the subsection heading. Use the \nameref{label} command to cite SI items in the text.
\subsection*{S1 Video}
\label{movie-MinD-density-pill}
{\bf Movie that shows the evolution of MinD density over a 250 second
  interval in a wild-type, pill-shaped cell.}  There is one frame for
every half second of simulation time. Above are results from the
stochastic model and below are results from the deterministic model.
The density shown is integrated over the direction normal to the
screen, so that the data is shown in molecules/$\micron^2$.

\subsection*{S2 Video}
\label{movie-MinD-density-shape-A}
{\bf Movie that shows the evolution of MinD density over a 350 second
  interval in the shape A cell.}  There is one frame for every half
second of simulation time. Above are results from the stochastic model
and below are results from the deterministic model.  The density shown
is integrated over the direction normal to the screen, so that the
data is shown in molecules/$\micron^2$.

\subsection*{S3 Video}
\label{movie-MinD-density-shape-B}
{\bf Movie that shows the evolution of MinD density over a 350 second
  interval in the shape B cell.}  There is one frame for every half
second of simulation time. Above are results from the stochastic model
and below are results from the deterministic model.  The density shown
is integrated over the direction normal to the screen, so that the
data is shown in molecules/$\micron^2$.

\subsection*{S4 Video}
\label{movie-MinD-density-stadium-A}
{\bf Movie that shows the evolution of MinD density over a 350 second
  interval in the stadium A cell.}  There is one frame for every half
second of simulation time. Above are results from the stochastic model
and below are results from the deterministic model.  The density shown
is integrated over the direction normal to the screen, so that the
data is shown in molecules/$\micron^2$.

\subsection*{S5 Video}
\label{movie-MinD-density-stadium-B}
{\bf Movie that shows the evolution of MinD density over a 350 second
  interval in the stadium B cell.}  There is one frame for every half
second of simulation time. Above are results from the stochastic model
and below are results from the deterministic model.  The density shown
is integrated over the direction normal to the screen, so that the
data is shown in molecules/$\micron^2$.

% Do NOT remove this, even if you are not including acknowledgments.
\section*{Acknowledgments}
Cras egestas velit mauris, eu mollis turpis pellentesque sit
amet. Interdum et malesuada fames ac ante ipsum primis in
faucibus. Nam id pretium nisi. Sed ac quam id nisi malesuada
congue. Sed interdum aliquet augue, at pellentesque quam rhoncus
vitae.

\nolinenumbers

% Compile your BiBTeX database using our plos2009.bst
% style file and paste the contents of your .bbl file
% here.

\bibliography{paper-MinD}

\end{document}