### 2015-06-11 paper/response-to-referees.txt
add response-to-referees.txt outlining what needs doing Tested everything on bennet.

Dear Editor,

We thank the referees for their helpful comments.  We have made a
number of revisions to our paper in response to their suggestions,
which we enumerate below, along with our response to each comment and
suggestion of the referees.  Each referee had several major concerns
and a few minor suggestions, which we will address in the order that
they were given, beginning with Referee #1.

Referee #1.

1. The largest concern of Referee #1 was our choice of Huang et al's
model for our simulations.  The referee acknowledges that we can stay
with this model, but feels that our justifaction that this model is
inadequate.  The referee objects that the fact that this model has
been widely is an inappropriate justification.  While we agree that
this alone is not a sufficient reason for studying this model, we do
believe that it is relevant: this is not some obscure older model that
we are testing, but rather the basis of all more recent models.

 FIXME NEED TO COME UP WITH MORE JUSTIFICATION? OR SIMPLY STATE THAT
 WE THINK IT IS RELEVANT?

2. As a secondary point, the referee points out we cannot conclude
that some other deterministic model could not demonstrate chaotic
behavior that explains the experimental results of Mannik et al.  We
agree that this cannot be ruled out, and have ADDED CLARIFICATION
WHERE?

3. The referee when looking at Figure 4 observes that the stadium
shapes exhibit osciallation that is largely bipolar, and expresses
doubt in our claim that the oscillations in stadium shapes are
irregular.  In particular, the referee points out that if we average
over many periods, we would find that the oscillations are indeed
bipolar.  We agree with this statement, but not with the conclusion.
Averaging the locations of the maxima would indeed miss the point of
the comparison with experiment, which is in examining the sequence of
locations of maxima.

We have EDITED WHAT? to moderate our claims.

Minor points:

4. The referee points out that MinD clusters had been considered prior
to Huang.  We have EDITED WHAT to include this reference.

5. ANOTHER EDIT TO DISCUSSION OF EARLIER WORK.

6. The referee points out that two previous works have also
demonstrated that stochastic simulations are needed in order to
explain WHAT?.  WE HAVE EDITED OUR DISCUSSION TO ILLUSTRATE THIS.

Referee #3.

7. The referee #3 also expressed doubt about our conclusion with
regard to the effect of irregular shape which we base on our stadium
models.  Unlike the previous referee, the criticism is based not on
the spatial output (Figure 4), but rather on the reported coherence
times.  We had attempted to make a distinction between spatially
irregular oscillation and temporally irregular oscillations.  This
distinction was evidently not clear, so we have ADDED AN EXPLICIT
DISCUSSION OF THE FACT THAT WE OBSERVE A GREATER DISRUPTION IN
TEMPORAL REGULARITY IN THE IRREGULARLY SHAPED CELLS.  FIXME WHAT DO WE
DO ABOUT SPATIAL IRREGULARITY?

8. The referee also questioned how the cell was divided in
order to perform the coherence time calculations, pointing out that a
smaller region could lead to a shorter coherence time.  EXPLAIN HOW WE
DIVIDE THE CELLS UP IN THIRDS BY VOLUME? IT'S AWKWARD THAT WE DON'T
DIVIDE ALONG LONG AXIS.

9. The referee points out that the "starry night" patterns are
inconsistent with experiments.  We agree with the referee that
membrane diffusion will alleviate this problem.  WE HAVE ADDED A
CLARIFICATION TO THIS EFFECT WHERE?

Minor points:

10. The referee requested a clarification to the caption of Figure 1,
which we HAVE ADDED.

11. The referee suggests that we should write $\mu m^2$ over $\mu^2$
for square microns.  We have made this change.  FIXME DO SO.

12. The referee suggested a title is needed in the caption of Figure
number 4.  WE HAVE ADDED A TITLE.

13. As requested WE HAVE CORRECTED THE CAPITALIZATION OF "E. coli".
FIXME.

14. The referee corrected us regarding the location at which the FtsZ
polymer forms, which is not on the cell wall, but rather near the
cytoplasmic side of the inner membrane.