### 2015-06-20 paper/response-to-referees.txt
more changes on Saturday

Dear Editor,

We thank the referees for their helpful comments.  We have made a
number of revisions to our paper in response to their suggestions,
which we enumerate below, along with our response to each comment and
suggestion of the referees.  Each referee had several major concerns
and a few minor suggestions, which we will address in the order that
they were given, beginning with Referee #1.

Referee #1.

1. The largest concern of Referee #1 was our choice of Huang et al's
model for our simulations.  The referee acknowledges that we can stay
with this model, but feels that our justifaction that this model is
inadequate.  The referee objects that the fact that this model has
been widely is an inappropriate justification.  While we agree that
this alone is not a sufficient reason for studying this model, we do
believe that it is relevant: this is not some obscure older model that
we are testing, but rather the basis of all more recent models.

 <Jeff gives this a quick shot!>

 FIXME NEED TO COME UP WITH MORE JUSTIFICATION? OR SIMPLY STATE THAT
 WE THINK IT IS RELEVANT?

2. As a secondary point, the referee points out we cannot conclude
that some other deterministic model could not demonstrate chaotic
behavior that explains the experimental results of Mannik et al.  We
agree that this cannot be ruled out, and have added mention of this to
our "Comparison of Experimental and Stadium Shapes", and have made
even more tentative our suggestion in the summary that other similar
deterministic models may also prove to be non-chaotic.

3. The referee when looking at Figure 4 observes that the stadium
shapes exhibit osciallation that is largely bipolar, and expresses
doubt in our claim that the oscillations in stadium shapes are
irregular.  In particular, the referee points out that if we average
over many periods, we would find that the oscillations are indeed
bipolar.  We agree with this statement, but not with the conclusion.
Averaging the locations of the maxima would indeed miss the point of
the comparison with experiment, which is in examining the sequence of
locations of maxima.

We have EDITED WHAT? to moderate our claims.

  <Jeff does this.>

Minor points:

4. The referee points out that MinD clusters had been considered by
Kruse prior to Huang.  We thank the referee for pointing out this
oversight, and have updated our background section to include this
reference and more accurately reflect the historical development.

5. Based on the referee's suggestion, we have clarified that reference
11 does not extend Huang's model, except in that it simulates the same
reaction-diffusion equations stochastically.

6. The referee points out that two previous works have also
demonstrated that stochastic simulations are needed in order to
explain WHAT?.  WE HAVE EDITED OUR DISCUSSION TO ILLUSTRATE THIS.

  <David does this.>

Referee #3.

7. The referee #3 also expressed doubt about our conclusion with
regard to the effect of irregular shape which we base on our stadium
models.  Unlike the previous referee, the criticism is based not on
the spatial output (Figure 4), but rather on the reported coherence
times.  We had attempted to make a distinction between spatially
irregular oscillation and temporally irregular oscillations.  This
distinction was evidently not clear, so we have ADDED AN EXPLICIT
DISCUSSION OF THE FACT THAT WE OBSERVE A GREATER DISRUPTION IN
TEMPORAL REGULARITY IN THE IRREGULARLY SHAPED CELLS.  JEFFSAYS{Changed
the last sentence before the conclusion.  I don't think it needs much
changing except maybe moderate our claims a bit (which I tried to
do).}  FIXME WHAT DO WE DO ABOUT SPATIAL IRREGULARITY?

  <Jeff does this.>

8. The referee also questioned how the cell was divided in order to
perform the coherence time calculations, pointing out that a smaller
region could lead to a shorter coherence time.  EXPLAIN HOW WE DIVIDE
THE CELLS UP IN THIRDS BY VOLUME? IT'S AWKWARD THAT WE DON'T DIVIDE
ALONG LONG AXIS.  JEFFSAYS{Bad news here.  It looks like I did this
dividing up the cells by eye.  Originally the slices were used to
created the box plots, which was only a way to visualize the data in
smaller time domains, so there's wasn't really any disadvantage to
doing it roughly like this (we even had the visually appealing little
colored diagrams on the sides of the box plots).  Not sure what to say
abou this in text.}

  <Jeff discusses this.  David thinks it was by thirds of volume.>

9. The referee points out that the "starry night" patterns are
inconsistent with experiments.  We agree with the referee that
membrane diffusion will alleviate this problem.  WE HAVE ADDED A
CLARIFICATION TO THIS EFFECT WHERE?

  <David>

Minor points:

10. The referee requested a clarification to the caption of Figure 1,
which we HAVE ADDED. JEFFSAYS{huh?}

  <Jeff>

11. The referee suggests that we should write $\mu m^2$ over $\mu^2$
for square microns.  We have made this change.  FIXME DO SO.

  <David>

12. The referee suggested a title is needed in the caption of Figure
number 4.  WE HAVE ADDED A TITLE.

  <Jeff>

13. As requested WE HAVE CORRECTED THE CAPITALIZATION OF "E. coli".
FIXME.

  <David>

14. The referee corrected us regarding the location at which the FtsZ
polymer forms, which is not on the cell wall, but rather near the
cytoplasmic side of the inner membrane.

  <David>