### 2015-08-01 paper/response-to-referees.txt
edits while meeting with Jeff and Rene

Dear Editor,

We thank the referees for their helpful comments.  We have made a
number of revisions to our paper in response to their suggestions,
which we enumerate below, along with our response to each comment and
suggestion of the referees.  Each referee had several major concerns
and a few minor suggestions, which we will address in the order that
they were given, beginning with Referee #1.

Referee #1.

1. The largest concern of Referee #1 was our choice of Huang et al's
model for our simulations.  The referee acknowledges that we can stay
with this model, but feels that our justifaction that this model is
inadequate.  The referee objects that the fact that this model has
been widely is an inappropriate justification.  While we agree that
this alone is not a sufficient reason for studying this model, we do
believe that it is relevant: this is not some obscure older model that
we are testing, but rather the basis of all more recent models.  We
believe that the limitations of this model are of interest, and have
added some discussion to the end of the introduction section
explaining this.

2. As a secondary point, the referee points out we cannot conclude
that some other deterministic model could not demonstrate chaotic
behavior that explains the experimental results of Mannik et al.  We
agree that this cannot be ruled out, and have added mention of this to
our "Comparison of Experimental and Stadium Shapes", and have made
even more tentative our suggestion in the summary that other similar
deterministic models may also prove to be non-chaotic.

3. The referee when looking at Figure 4 observes that the stadium
shapes exhibit oscillation that is largely bipolar, and expresses
doubt in our claim that the oscillations in stadium shapes are
irregular.  In particular, the referee points out that if we average
over many periods, we would find that the oscillations are indeed
bipolar.  We agree with this statement, but not with the conclusion.
Averaging the locations of the maxima would indeed miss the point of
the comparison with experiment, which is in examining the sequence of
locations of maxima.

We have edited the "comparison of experimental and stadium shapes"
section, the conclusion, and the abstract to moderate our claims
regarding the spatial irregularity of the experimental and stadium
shapes as predicted by the stochastic model.

Minor points:

4. The referee points out that MinD clusters had been considered by
Kruse prior to Huang.  We thank the referee for pointing out this
oversight, and have updated our background section to include this
reference and more accurately reflect the historical development.

5. Based on the referee's suggestion, we have clarified that reference
11 does not extend Huang's model, except in that it simulates the same
reaction-diffusion equations stochastically.

6. The referee points out that we failed to mention that stochastic
simulations are needed to explain the experimentally observed
length-dependence of fluctuations in the Min system.  We have added
mention of this additional evidence of the importance of inclusion of
stochastic effects, in the paragraph suggested by the referee.

Referee #3.

7. The referee #3 also expressed doubt about our conclusion with
regard to the effect of irregular shape which we base on our stadium
models.  Unlike the previous referee, the criticism is based not on
the spatial output (Figure 4), but rather on the reported coherence
times.  We agree with the referee that our concluding sentence
immediately prior to the conclusion section was unjustified, and have
edited this sentence to reflect the referee's objections.

8. The referee also questioned how the cell was divided in order to
perform the coherence time calculations, pointing out that a smaller
region could lead to a shorter coherence time.  We have added an
explanation of how we divided the cell into three sections, and gave
numbers showing the relative volume of the sections.  These were not
identical, and thus will have some effect on the correlation times
computed.  However, the region volumes are roughly similar, so we do
not expect this effect to be dramatic.

9. The referee points out that the "starry night" patterns are
inconsistent with experiments.  We agree with the referee that
membrane diffusion will alleviate this problem.  We have added a
clarification that this effect is *not* observed in experiment.

Minor points:

10. As the referee requested, we have clarified the caption of Figure
1, to include the fact that these reactions are for a Min system
model that is used in this work.

11. The referee suggests that we should consistently write $\mu m^2$
over $\mu^2$ for square microns.  We have made this change to the plot
that used $\mu^2$.

12. The referee suggested a title is needed in the caption of Figure
number 4.  We have edited the caption to begin with an appropriate
title-like sentence fragment, rather than "We display here...".

13. As requested we have corrected the capitalization of "E. coli".

14. The referee corrected us regarding the location at which the FtsZ
polymer forms, which is not on the cell wall, but rather near the
cytoplasmic side of the inner membrane.  We have corrected this
statement in the first paragraph of the introduction.